US Korea Exchange Rates \\ STA206 Final Project

### 2014-12-14 paper.tex
centered model

We analyze data on exchange rate between the US and Korea from 1999 to
2014. In 1997 the IMF crisis in Korea perturbed the economic indicators,
but in looking at the data we can see that it has been stable since 1999.

@@

\begin{itemize}
    \item {\tt gdp} The gross domestic product. Units: USD Million
    \item {\tt unemployment rate} The percentage of the labor force who are unemployed and actively seeking work. Units: Percent (monthly)
    \item {\tt exports} The total value of the goods and services
produced domestically and purchased by foreign entities. Units: USD Million (Monthly)
    \item {\tt imports} The total value of a country's imports of physical goods and payments to foreigners for services like shipping and tourism. Units: USD Million (Monthly)
    \item {\tt interest rate} The monthly average of the central bank policy rate. This is the interest rate the central bank charges on loans to commercial banks. Units: Percent (Monthly)
    \item {\tt inflation rate} The growth rate of the prices. (Monthly)
    \item {\tt consumer produce index} The Consumer Price Index (CPI) is a measure of inflation related to the cost of living. Units: Index Points 2010=100, NSA (Monthly)
    \item {\tt debt} The total amount of public and private debt owned by foreign creditors. Units: USD Million Current Prices, NSA (quarterly)
    \item {\tt gdp deflator} The relative difference between the real and nominal GDPs. Units: Index Points NSA
    \item {\tt goverment spending} The yearly expenditure of the federal
        government. Units: local currency
    \item {\tt political party} Categorical variable for the political
party of the president.
\end{itemize}

@@

\subsection{Reproducibility}

@@

Modern high level programming languages like R provide incredible tools. By
using their capabilities we can

We started out with some general summary statistics and plotting. The first
issue was missing data, which we had anticipated. While most of the
variables were reported monthly, \texttt{debt} was reported on a quarterly
basis and 
Our simple approach for dealing with this was to
make a plot of these variables over time.
If they are approximately linear then we could 
linearly interpolate with respect to time. The scaled plots 
Figure~\ref{fig:na_plot_KOR} and
Figure~\ref{fig:na_plot_USA} do appear to be mostly linear, so we interpolated. 
In this plots one can observe that \texttt{debt\_USA} is decreasing; this is
because \texttt{debt} carries a sign. The value of \texttt{debt\_USA} on

Because exchange rate is a ratio of Korean currency / US dollar we decided
to make transformed variables consisting of the Korean economic indicator
divided by the corresponding US indicator. We also included
\texttt{inflation} untransformed after some preliminary analysis 
indicated that it was
important. \texttt{Date} was intentionally left out of the model in order
to be able to interpret the model in terms of the other parametrs. However,
the effects of time are still present through the variables that are linear
over time.

The variables under further investigation are:
\texttt{gdp}, \texttt{interest\_rate}, \texttt{inflation\_KOR}, 
\texttt{inflation\_USA}, and \texttt{govspending}.

The box - cox transformation in figure~\ref{fig:boxcox} suggests a
transformation of $1 / x^2$, but this extreme and is due to the outliers.
We used the multiplicative inverse transformation and examined the
residual plots figure~\ref{fig:cooks} to see why this is the case.

We see
that the points 113 and 117 are the primary outlying cases. Looking up the
corresponding dates reveals that these are the points occuring during
October 2008 and February 2009, during the subprime mortgage crisis
which shook the global economy. There are no points between 113 and 117
because the test train split didn't select them. In the validation set we
discovered similar results for this period. The subprime mortgage crisis is
the reason that they are outliers.

Attempting to reference \cite{lamport94}.

\newpage
\section{Appendix}

Model Selection Output:
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & sse & bic & r2 & r2a & cp \\
  \hline
1 & 0.00 & -46.47 & 0.37 & 0.37 & 411.04 \\
  2 & 0.00 & -178.62 & 0.80 & 0.79 & 55.28 \\
  3 & 0.00 & -185.95 & 0.82 & 0.81 & 41.04 \\
  4 & 0.00 & -206.51 & 0.85 & 0.85 & 13.91 \\
  5 & 0.00 & -212.36 & 0.86 & 0.86 & 5.39 \\
  6 & 0.00 & -210.43 & 0.87 & 0.86 & 4.72 \\
  7 & 0.00 & -207.14 & 0.87 & 0.86 & 5.33 \\
  8 & 0.00 & -202.47 & 0.87 & 0.86 & 7.22 \\
  9 & 0.00 & -197.85 & 0.87 & 0.86 & 9.06 \\
  10 & 0.00 & -193.12 & 0.87 & 0.86 & 11.00 \\
   \hline
\end{tabular}
\end{table}

@@

\begin{figure}
  \centering
    \includegraphics{na_plot_USA.pdf}
  \caption{Scaled variables with NA values}
  \label{fig:na_plot_USA}
\end{figure}

\begin{figure}
  \centering
    \includegraphics{scatterplot.pdf}
  \caption{Scatterplot of untransformed data}
  \label{fig:scatterplot}
\end{figure}

@@

@@

\begin{thebibliography}{9}

\bibitem{lamport94}
  Leslie Lamport,
  \emph{\LaTeX: a document preparation system}.
  Addison Wesley, Massachusetts,
  2nd edition,
  1994.

\end{thebibliography}

>>>>

We analyze data on exchange rate between the US and Korea from 1999 to
2014. In 1997 the IMF crisis in Korea perturbed the economic indicators,
but they have been mostly stable since 1999.

@@

\begin{itemize}
    \item {\tt gdp} The gross domestic product. Units: USD Million
    \item {\tt unemployment rate} The percentage of the labor force who are unemployed and actively seeking work. Units: Percent (monthly)
    \item {\tt exports} The total value of the goods and services
produced domestically and purchased by foreign entities. Units: USD Million (Monthly)
    \item {\tt imports} The total value of a country's imports of physical goods and payments to foreigners for services like shipping and tourism. Units: USD Million (Monthly)
    \item {\tt interest rate} The monthly average of the central bank policy rate. This is the interest rate the central bank charges on loans to commercial banks. Units: Percent (Monthly)
    \item {\tt inflation rate} The growth rate of the prices. (Monthly)
    \item {\tt consumer produce index} The Consumer Price Index (CPI) is a measure of inflation related to the cost of living. Units: Index Points 2010=100, NSA (Monthly)
    \item {\tt debt} The total amount of public and private debt owned by foreign creditors. Units: USD Million Current Prices, NSA (quarterly)
    \item {\tt gdp deflator} The relative difference between the real and nominal GDPs. Units: Index Points NSA
    \item {\tt goverment spending} The yearly expenditure of the federal
        government. Units: local currency
\end{itemize}

@@

Through a little engineering the data analysis process
can be made transparent, automated, extensible, and fully reproducible.
For more information on the techniques please refer to the section on
Reproducibility in the appendix.

We started out with some general summary statistics and plotting. The first
issue was missing data, which we had anticipated. While most of the
variables were reported monthly, some were reported on a quarterly or
annual basis. (Figure~\ref{fig:na_plot_KOR})
Our simple approach for dealing with this was to
make a plot of these variables over time.
If they are approximately linear then we could 
linearly interpolate with respect to time. The scaled plots 
Figure~\ref{fig:na_plot_KOR} and
Figure~\ref{fig:na_plot_USA} do appear to be mostly linear, so we interpolated. 
In this plots one can observe that \texttt{debt\_USA} is decreasing; this is
because \texttt{debt} carries a sign.

Because exchange rate is a ratio of Korean currency / US dollar we decided
to make transformed variables consisting of the Korean economic indicator
divided by the corresponding US indicator. We also included
\texttt{inflation} untransformed since the US has value 0 for some time
periods, meaning that this transformation is not defined.
\texttt{Date} was intentionally left out of the model in order
to be able to interpret the model in terms of the other parametrs. However,
the effects of time are still present through the variables that are linear
over time.

The variables under further investigation are:
\texttt{gdp}, \texttt{interest\_rate}, \texttt{inflation\_KOR}, 
\texttt{inflation\_USA}, and \texttt{govspending}. (Model 1)

The box - cox transformation in figure~\ref{fig:boxcox} suggests a
transformation of $1 / x^2$, but this is extreme and is due to the
outliers. Instead we used the multiplicative inverse transformation on the
\texttt{exchange\_rate}, which has the appealing property of maintaining
interpretability; instead of Korean Won / USD we have USD / Korean Won.

The outliers can be readily detected in the diagnostic plots 
figure~\ref{fig:cooks}. We see
that the points 113 and 117 are the primary outlying cases, which have
absolute DFFIT values of 1.16 and 0.89. This is much bigger than the cutoff
of $2 / \sqrt{p / n} = 0.445$. Cook's distance is large as well.
Looking up the
corresponding dates reveals that these are the points occuring during
October 2008 and February 2009, during the subprime mortgage crisis
which shook the global economy. There are no points between 113 and 117
because the test train split didn't select them. In the validation set we
discovered similar results for this period(114, 118). The subprime mortgage
crisis is the reason that they are outliers. So, we took off the outliers
from our data set to get model 1 fit~\ref{m1:train}. We also had trouble
with recent dates having large leverage values of around 0.16, which is
greater than the threshold of $2 * p / n = 0.08$.

When checking this final model with the validation set, then the
\texttt{govspending} variable is not
significant (P-value:0.312)~\ref{m1:test}. The added-variable plot in
figure~\ref{fig:added} indicates \texttt{govspending} contains no
additional information useful for predicting (1/exchage\_rate) beyond the
other variable, so it is not helpful to add \texttt{govspending} to our
model. Thus, we decide the our final variables are \texttt{gdp},
\texttt{interest\_rate}, \texttt{inflation\_KOR}, and
\texttt{inflation\_USA}~\ref{finalmodel}. The corresponding equation is:

$\frac{1}{exchange\_rate} = 3.868323e^{-4} + 9.692157e^{-3} * gdp
-1.289123e^{-5} *interest\_rate -2.573026e^{-5} * inflation\_KOR +
2.135945e^{-5}  * inflation\_USA$

For this model we have $R^2 = 0.85, MSE = 1.28^{-9}$.

\subsection{Centering and Signs}

The model above is difficult to interpret, so we decided to center the data
to minimize multicollinearity and get more accessible numbers. After a
similar analysis process as described above we found that the final model
for the centered data has the same variables as well as \texttt{export}.

common variables and export variable. Interesting part of this trial is
about outliers. Its outliers are showing up 2000 finanical crisis rather
than the subprime mortage crisis (it fitted well)~\ref{centered}.

We used Bonferroni's method to construct simultaneous 95\% confidence intervals
for each of the 5 variables selected in the centered model.

Our model: $\frac{USA dollar Worthy}{KOREA Won Worthy} \sim \frac{KOREA gdp}{USA GDP} - \frac{KOREA Interest Rate}{USA Interest Rate} -  Korea inflation + USA inflation$, means when relative Korea gdp over US gdp is increasing, the won per a dollar is decreasing. The Korea Interest Rate is relatively increased over the USA Interest rate(higher interest rate means higher liquidity of money in markets), which bring the won per a dollar is increasing. Also, the signs of Korea inflation and USA inflation are different, which is very reliable because we are predicting the $\frac{USA dollar Worthy}{KOREA Won Worthy}$.

Commonly, we can think if you want to exchage your currency to other foriegn currency, it would better off to do it when your country's relative gdp is greater than the foriegn, when your country's relative interest rate is less , when your contry is not on inflation, or when the foriegn country is on inflation situation.
We have noticed exchange rates are dependent on previous exchage rates(time series) if we are available to do mutltivariate time series.

\section{Appendix - Reproducibility}

@@

\section{Appendix 1}

@@

\begin{figure}
  \centering
    \includegraphics{na_plot_USA.pdf}
  \caption{This figure illustrates the perils of depending on live data. US debt was
formerly measured as a negative number, which is why it was decreasing. But
midway through our analysis the source began to return a huge positive
number for the most recent dates.}

  \label{fig:na_plot_USA}
\end{figure}
\begin{figure}
  \centering
    \includegraphics{scatterplot.pdf}
  \caption{Scatterplot of untransformed data}
  \label{fig:scatterplot}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics{Addedvariable.pdf}
	\caption{Added-Variable Plot for Government Spending}
	\label{fig:added}
\end{figure}

@@

\section{Appendix 2}

\subsection{Model Selection Output}
\begin{table}[ht]
	\centering
	\begin{tabular}{rrrrrr}
		\hline
		& sse & bic & r2 & r2a & cp \\
		\hline
		1 & 0.00 & -46.47 & 0.37 & 0.37 & 411.04 \\
		2 & 0.00 & -178.62 & 0.80 & 0.79 & 55.28 \\
		3 & 0.00 & -185.95 & 0.82 & 0.81 & 41.04 \\
		4 & 0.00 & -206.51 & 0.85 & 0.85 & 13.91 \\
		5 & 0.00 & -212.36 & 0.86 & 0.86 & 5.39 \\
		6 & 0.00 & -210.43 & 0.87 & 0.86 & 4.72 \\
		7 & 0.00 & -207.14 & 0.87 & 0.86 & 5.33 \\
		8 & 0.00 & -202.47 & 0.87 & 0.86 & 7.22 \\
		9 & 0.00 & -197.85 & 0.87 & 0.86 & 9.06 \\
		10 & 0.00 & -193.12 & 0.87 & 0.86 & 11.00 \\
		\hline
	\end{tabular}
\end{table}

\subsection{Influence Measures}

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrrrrr}
  \hline
 & dfb.1\_ & dfb.gdp & dfb.int\_ & dfb.i\_KO & dfb.i\_US & dfb.gvsp & dffit & cov.r & cook.d & hat \\
  \hline
110 & -0.15 & -0.04 & -0.09 & 0.18 & 0.06 & 0.11 & 0.25 & 1.16 & 0.01 & 0.12 \\
  113 & 0.67 & 0.18 & 0.43 & -0.93 & 0.00 & -0.51 & -1.17 & 0.30 & 0.18 & 0.05 \\
  117 & 0.24 & 0.01 & 0.21 & -0.65 & 0.70 & -0.18 & -0.89 & 0.67 & 0.12 & 0.07 \\
  122 & -0.02 & -0.02 & 0.01 & -0.00 & 0.07 & 0.01 & -0.08 & 1.18 & 0.00 & 0.11 \\
  173 & -0.03 & -0.03 & -0.03 & -0.00 & -0.01 & 0.05 & 0.08 & 1.16 & 0.00 & 0.09 \\
  176 & -0.05 & -0.05 & -0.05 & 0.01 & 0.00 & 0.09 & 0.11 & 1.24 & 0.00 & 0.16 \\
  177 & -0.06 & -0.05 & -0.05 & 0.01 & -0.01 & 0.09 & 0.11 & 1.24 & 0.00 & 0.15 \\
  178 & -0.06 & -0.05 & -0.05 & 0.02 & -0.00 & 0.10 & 0.11 & 1.24 & 0.00 & 0.16 \\
  179 & -0.21 & -0.20 & -0.18 & 0.06 & 0.02 & 0.33 & 0.38 & 1.20 & 0.02 & 0.16 \\
  180 & -0.37 & -0.33 & -0.30 & 0.13 & 0.04 & 0.57 & 0.64 & 1.12 & 0.07 & 0.16 \\
   \hline
\end{tabular}
\end{table}

\subsubsection{Model1 summary output (Train Data)\label{m1:train}}
\begin{verbatim}
Call:
lm(formula = I(1/exchange_rate) ~ ., data = tdata8)

Residuals:
Min         1Q     Median         3Q        Max 
-7.807e-05 -1.661e-05 -3.530e-07  2.128e-05  6.973e-05

Coefficients:
               Estimate   Std. Error t value Pr(>|t|)    
(Intercept)    3.901e-04  2.851e-05  13.683  < 2e-16 ***
gdp            1.130e-02  5.522e-04  20.471  < 2e-16 ***
interest_rate -1.211e-05  1.085e-06 -11.158  < 2e-16 ***
inflation_KOR -1.877e-05  3.334e-06  -5.630 1.35e-07 ***
inflation_USA  1.529e-05  2.574e-06   5.943 3.23e-08 ***
govspending   -9.170e-06  3.423e-06  -2.679   0.0085 ** 
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 2.863e-05 on 112 degrees of freedom
Multiple R-squared:  0.9034,	Adjusted R-squared:  0.8991 
F-statistic: 209.4 on 5 and 112 DF,  p-value: < 2.2e-16
\end{verbatim}

\subsubsection{Model1 summary output (Validation Data)\label{m1:test}}

\begin{verbatim}
Call:
lm(formula = I(1/exchange_rate) ~ ., data = test99)

Residuals:
Min         1Q     Median         3Q        Max 
-6.962e-05 -1.878e-05 -2.636e-06  1.988e-05  4.937e-05

Coefficients:
               Estimate   Std. Error t value Pr(>|t|)    
(Intercept)    4.038e-04  3.592e-05  11.244 1.56e-15 ***
gdp            1.001e-02  7.838e-04  12.769  < 2e-16 ***
interest_rate -1.163e-05  1.601e-06  -7.268 1.85e-09 ***
inflation_KOR -2.808e-05  3.749e-06  -7.492 8.15e-10 ***
inflation_USA  2.315e-05  4.847e-06   4.776 1.50e-05 ***
govspending   -3.640e-06  4.246e-06  -0.857    0.395    
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 2.742e-05 on 52 degrees of freedom
Multiple R-squared:  0.9161,	Adjusted R-squared:  0.908 
F-statistic: 113.5 on 5 and 52 DF,  p-value: < 2.2e-16
\end{verbatim}

\subsubsection{Summary table of Final model}

\label{finalmodel}
\begin{verbatim}
Call:
lm(formula = I(1/exchange_rate) ~ gdp + interest_rate + inflation_KOR + 
inflation_USA, data = alldata)

Residuals:
Min         1Q     Median         3Q        Max 
-1.631e-04 -1.924e-05  4.079e-06  2.177e-05  8.537e-05

Coefficients:
               Estimate   Std. Error t value Pr(>|t|)    
(Intercept)    3.868e-04  2.285e-05  16.929  < 2e-16 ***
gdp            9.692e-03  4.007e-04  24.190  < 2e-16 ***
interest_rate -1.289e-05  8.636e-07 -14.927  < 2e-16 ***
inflation_KOR -2.573e-05  2.637e-06  -9.758  < 2e-16 ***
inflation_USA  2.136e-05  2.646e-06   8.072 1.04e-13 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 3.58e-05 on 176 degrees of freedom
Multiple R-squared:  0.8555,	Adjusted R-squared:  0.8522 
F-statistic: 260.5 on 4 and 176 DF,  p-value: < 2.2e-16
\end{verbatim}

\subsubsection{Final model Summary Output(Centered data)}
\label{centered}

\begin{verbatim}
Call:
lm(formula = exchange_rate ~ gdp + exports + interest_rate + 
inflation_KOR + inflation_USA, data = sdata_n)

Residuals:
Min       1Q   Median       3Q      Max 
-0.72512 -0.22635  0.00475  0.16783  0.94453

Coefficients:
              Estimate    Std. Error t value Pr(>|t|)    
(Intercept)   -0.03923    0.02417  -1.623    0.106    
gdp           -0.68565    0.04872 -14.073  < 2e-16 ***
exports       -0.24752    0.05083  -4.869 2.54e-06 ***
interest_rate  0.59072    0.03611  16.361  < 2e-16 ***
inflation_KOR  0.28550    0.02762  10.337  < 2e-16 ***
inflation_USA -0.29924    0.03197  -9.361  < 2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 0.3211 on 171 degrees of freedom
Multiple R-squared:  0.8868,	Adjusted R-squared:  0.8835 
F-statistic: 267.9 on 5 and 171 DF,  p-value: < 2.2e-16
\end{verbatim}

\subsubsection{Bonferroni's confidence intervals on centered coefficients}

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & lower & upper \\
  \hline
(Intercept) & -0.10 & 0.02 \\
  gdp & -0.80 & -0.57 \\
  exports & -0.37 & -0.13 \\
  interest\_rate & 0.51 & 0.68 \\
  inflation\_KOR & 0.22 & 0.35 \\
  inflation\_USA & -0.37 & -0.22 \\
   \hline
\end{tabular}
\end{table}
